---
title: "An offline streaming workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An offline streaming workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

PhysioStream governs *live* physiological data streams: Lab Streaming Layer
inlets and outlets, XDF interchange, MQTT/TTL stimulation triggers, and a live
biofeedback scope. Those transports need a device or a running network and are
not exercised here.

Everything else runs **in process, with no device**. This vignette walks the
device-free path end to end on a tiny synthetic signal: stream metadata, the
deterministic loopback source, a snapshot to a `PhysioExperiment`, a synchronous
processing pipeline, online DSP, clock correction, and a governed stimulation
trigger. Every code block below is executed when the vignette is built.

```{r setup}
library(PhysioStream)
```

## Stream metadata

A `StreamInfo` is validated, immutable, and serializable. It is the single
source of truth for channel identity, rate, and storage type.

```{r info}
info <- streamInfo(
  "demo-eeg",
  type = "EEG",
  channel_names = c("C3", "C4"),
  nominal_srate = 100,
  channel_units = c("uV", "uV")
)
info
streamChannels(info)
```

## A deterministic loopback stream

`loopbackSource()` is an in-process `StreamSource` backed by a bounded native
ring buffer. It exercises the same lifecycle (`created` -> `open` -> `closed`)
and the same sample/timestamp governance as a live transport, but you feed it
yourself with `loopbackFeed()`.

```{r loopback}
source <- streamOpen(loopbackSource(info, capacity = 256L))
streamState(source)

n <- 50
t <- seq_len(n) / 100
samples <- cbind(
  C3 = sin(2 * pi * 10 * t),
  C4 = sin(2 * pi * 10 * t + 0.5)
)
loopbackFeed(source, samples, t)
```

## Snapshot to a PhysioExperiment

`streamSnapshot()` copies buffered samples into a `PhysioExperiment`, preserving
the original timestamps and the monotonic stream sequence. From here the rest of
the ecosystem applies.

```{r snapshot}
pe <- streamSnapshot(source)
dim(pe)
SummarizedExperiment::assayNames(pe)
```

## A synchronous processing pipeline

A `StreamPipeline` owns a bounded chunk queue and an ordered, causal operation
graph. Here we register the compiled second-order-section + rolling-RMS
operation. A pass-through SOS keeps the example self-contained; use
`PhysioPreprocess::sosDesign()` for a real band definition.

```{r pipeline}
sos <- matrix(
  c(1, 0, 0, 1, 0, 0), 1L, 6L,
  dimnames = list(NULL, c("b0", "b1", "b2", "a0", "a1", "a2"))
)
pipeline <- streamPipeline(chunk_size = 64L)
onChunk(pipeline, bandpassRmsOp(sos, window_samples = 4L))

pipelineEnqueue(pipeline, samples, t, ingest_time_ns = 0)
result <- pipelineStep(pipeline)
head(result$results[[1]]$output$samples)
```

## Online DSP

The DSP processors are stateful and chunk-invariant. An adaptive filter, for
example, estimates the part of a channel that is linearly predictable from a
paired reference.

```{r adaptive}
set.seed(1)
reference <- sin(2 * pi * 10 * t)
observed <- 0.4 * reference + rnorm(n, sd = 0.05)
filt <- lmsFilter(n_taps = 4L, step_size = 0.1)
residual <- update(
  filt,
  matrix(observed, ncol = 1),
  reference = matrix(reference, ncol = 1)
)
length(residual$output)
```

## Clock correction

Explicit clock models correct offset and drift and regularize jitter. They
change timestamps only -- samples are never interpolated or reordered.

```{r clock}
device <- seq(0, 10, by = 0.5)
offset <- -0.4 + 0.005 * (device - median(device))
model <- clockOffset(device, offset, method = "ols")
model$segments$drift_ppm

jittered <- c(10.001, 10.010, 10.021, 10.029)
as.numeric(dejitter(jittered, nominal_srate = 100))
```

## A governed stimulation trigger

The loopback trigger records commands and timing with no device, while
exercising the same arm, dead-man, maximum, duplicate-ID, and refractory
interlocks as the MQTT and TTL transports. Software interlocks are safeguards
only; independently fail-safe stimulation hardware is still required.

```{r trigger}
trigger <- loopbackTrigger(
  allowed_channels = "left",
  max_intensity = 20,
  intensity_unit = "mA",
  max_duration_ms = 500,
  refractory_ms = 0,
  deadman_ms = 1000
)
triggerOpen(trigger)
armTrigger(trigger, "demo-session", now_ns = 0)
receipt <- sendStim(
  trigger, intensity = 2.5, channel = "left",
  duration_ms = 125, command_id = "cmd-1", now_ns = 10
)
receipt$status
triggerClose(trigger)
```

## Where to go next

- **Live acquisition.** `lslResolveStreams()`, `lslInlet()`, and `lslOutlet()`
  bind Lab Streaming Layer streams; `lslAvailable()` reports whether the backend
  is installed.
- **File interchange.** `readXDF()` / `writeXDF()` require the pyxdf backend
  (`xdfAvailable()`).
- **Closed-loop control.** `closedLoop()` binds a detector
  (`emgOnsetOp()`, `erdIntentOp()`, `phaseTargetOp()`) to a trigger with
  commit-before-stimulation semantics.
- **Live visualization.** `biofeedbackScope()` provides a bounded display-only
  scope (Shiny is required only to launch it).
